\appendix
\etocdepthtag.toc{appendix} % 여기서부터 부록 시작임을 태그 지정
\section*{\Large Appendix}

% \setcounter{figure}{0}
% \setcounter{table}{0}
% \renewcommand{\thefigure}{A.\arabic{figure}}
% \renewcommand{\thetable}{A.\arabic{table}}

%%%%%%%%%%%% TABLE OF CONTENTS %%%%%%%%%%%% 
\makeatletter
\newcommand{\appendix@noopaddcontentsline}[3]{}
\let\appendix@origsection\section
\renewcommand{\section}{\@ifstar\appendix@origsection\appendix@section}
\newcommand{\appendix@section}[1]{%
  \let\appendix@savedaddcontentsline\addcontentsline
  \let\addcontentsline\appendix@noopaddcontentsline
  \appendix@origsection{#1}%
  \let\addcontentsline\appendix@savedaddcontentsline
  \phantomsection
  \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
}
\let\appendix@origsubsection\subsection
\renewcommand{\subsection}{\@ifstar\appendix@origsubsection\appendix@subsection}
\newcommand{\appendix@subsection}[1]{%
  \let\appendix@savedaddcontentsline\addcontentsline
  \let\addcontentsline\appendix@noopaddcontentsline
  \appendix@origsubsection{#1}%
  \let\addcontentsline\appendix@savedaddcontentsline
  \phantomsection
  \addcontentsline{toc}{subsection}{\protect\numberline{\thesubsection}#1}%
}
\makeatother

\etocsetlevel{author}{6}
\etocsetlevel{title}{6}
\etocsettagdepth{main}{none}
\etocsettagdepth{appendix}{subsection}
\makeatletter
{%
\let\section\appendix@origsection
\let\subsection\appendix@origsubsection
\tableofcontents
}
\makeatother

\etocdepthtag.toc{appendix} % 여기서부터 부록 시작임을 태그 지정
%%%%%%%%%%%% TABLE OF CONTENTS %%%%%%%%%%%% 



\clearpage
\section{Training Details}
\label{sec:appendix_experimental_details}
% Our primary interest is in transferability, \textit{i.e.}, whether learning a task in one modality improves performance on the corresponding task in the other modality. Thus, for the transferability experiments, our default strategy is to apply LoRA to the Query, Key, Value projection and MLP layers of the transformer blocks, which is  shared across modalities. All other major components, including the image encoder, projector/aligner, language modeling head, and diffusion head, are kept frozen. Across all runs, we use LoRA rank 128, 8-bit AdamW, a global batch size of 128, and bfloat16 training. To enable a fair comparison across architectures, we select models in the 7B--8B parameter range. For quantitative results, we report the best checkpoint from each training run.
% For BAGEL~\cite{deng2025bagel}, we use a learning rate of $1\times10^{-4}$ and retain the original dynamic image resizing strategy of the model. For Janus-Pro~\cite{chen2025januspro}, we use the Janus-Pro-7B checkpoint with a learning rate of $4\times10^{-5}$; for both understanding and generation, we preserve the original aspect ratio, resize the longer side to 384 pixels, and pad the shorter side to obtain a $384 \times 384$ input. For Lumina-DiMOO, we use a learning rate of $3\times10^{-6}$; for both understanding and generation, we preserve the original aspect ratio, resize the longer side to 512 pixels, and pad the shorter side to obtain a $512 \times 512$ input. For BLIP3-o~\cite{chen2505blip3}, we use the BLIP3o-8B checkpoint. Since BLIP3-o employs separate transformers for understanding and generation, we fine-tune the corresponding branch for each task. We retain the original image processing strategy of Qwen2.5-VL-7B-Instruct~\cite{bai2025qwen2}, which is used as LLM backbone of BLIP3-o.

In the experiments of Section~\ref{}, we evaluate four representative UMMs spanning different architectural families (Figure~\ref{fig:arch_categorize}): BAGEL~\cite{deng2025bagel} (7B), Janus-Pro~\cite{chen2025januspro} (7B), Lumina-DiMOO (?B), and BLIP3-o~\cite{chen2505blip3} (?B). All models fall within the 7B–8B parameter range to enable a fair comparison across architectures. To investigate whether transferability emerges in the shared LLM backbone, we perform LoRA-based supervised fine-tuning on top of publicly available pretrained weights. We apply LoRA with rank 128 to the query, key, and value projection layers as well as the MLP layers of the transformer blocks. All other components—including the image encoder, projector/aligner, language modeling head, and diffusion head—are kept frozen. We use 8-bit AdamW with bfloat16 training across all runs. Unless otherwise specified, we follow the official codebases and recommended training configurations of each model. The model-specific settings are described below.

\noindent\textbf{BAGEL~\cite{deng2025bagel}.}
We use a learning rate of $1\times10^{-4}$ and retain the model’s original dynamic image resizing strategy.

\noindent\textbf{Janus-Pro~\cite{chen2025januspro}.}
We use a learning rate of $4\times10^{-5}$. For both understanding and generation, we preserve the aspect ratio, resize the longer side to 384 pixels, and pad the shorter side, resulting in a $384\times384$ input.

\noindent\textbf{Lumina-DiMOO~\cite{xin2025lumina}.}
We use a learning rate of $3\times10^{-6}$. For both understanding and generation, we preserve the aspect ratio, resize the longer side to 512 pixels, and pad the shorter side, resulting in a $512\times512$ input. We use the same training configuration for the experiments in Sections~\ref{sec:tasks} and~\ref{sec:discuss_ocr}.

\noindent\textbf{BLIP3-o~\cite{chen2505blip3}.}
Since BLIP3-o employs separate transformer branches for understanding and generation, we fine-tune only the branch corresponding to each task. We retain the original image processing pipeline of Qwen2.5-VL-7B-Instruct~\cite{bai2025qwen2}, which serves as the LLM backbone of BLIP3-o.

% \begin{table}[h]
% \centering
% \caption{Model-specific training hyperparameters.}
% \label{tab:hyperparams}
% \setlength{\tabcolsep}{6pt}
% \begin{tabular}{lcccc}
% \toprule
% Hyperparameter & BAGEL & Janus-Pro & Lumina-DiMOO & BLIP3-o \\
% \midrule
% Learning rate          & $1 \times 10^{-4}$ & $4 \times 10^{-5}$ & $3 \times 10^{-6}$ & -- \\
% LR scheduler           & --                 & --                 & --                 & -- \\
% Weight decay           & --                 & --                 & --                 & -- \\
% Gradient clipping norm & --                 & --                 & --                 & -- \\
% Warm-up steps          & --                 & --                 & --                 & -- \\
% Steps                  & --                 & --                 & --                 & -- \\
% Batch size             & --                 & --                 & --                 & -- \\
% Input resolution       & --            & $384 \times 384$   & $512 \times 512$   & -- \\
% \bottomrule
% \end{tabular}
% \end{table}

\clearpage
\section{Task Details}
\label{sec:appendix_dataset_construction}
\subsection{Counting}
\subsubsection{Dataset detail.} Our \textit{visual instruction tuning dataset} for counting consists of 65.6K images and approximately 200K question-answer pairs. We derive the VQA pairs from the training sets of PixMo-Count~\cite{deitke2025molmo} and PixMo-Points~\cite{deitke2025molmo}.
For counting understanding, we directly adopt the original question-answer pairs from PixMo-Count, as it is originally designed for counting tasks. To scale up the training data, we additionally incorporate samples from PixMo-Points. Since PixMo-Points is a point prediction dataset rather than a counting dataset, we reformulate it for counting by treating the number of points to be predicted as the object count. Note that PixMo-Points originally contains annotations more thank 1K points per image. We find that such extreme cases introduce excessive difficulty during training. To maintain a reasonable difficulty level, we filter out samples whose object count exceeds 20. For \textit{counting-aware generation dataset}, we construct the training set from the same 65.6K images used in the understanding split to ensure a controlled experimental setting. Specifically, we leverage the object count annotations from the understanding training set and use Qwen3-VL~\cite{bai2025qwen3} to generate count-aware generation prompts conditioned on these annotations.
Table~\ref{tab:dataset_counting} presents representative samples from both the counting understanding and generation training sets.
\input{x_supple/tabs/dataset_counting}


% \subsubsection{Understanding evaluation.}
% We evaluate counting understanding using 540 QA pairs constructed from the test split of PixMo-Count. Each QA pair explicitly includes both the question and a response template. In particular, the prompt contains a \texttt{Response Example} field that asks the model to place the predicted count inside the markdown markers \texttt{** **}. During evaluation, we first parse the substring enclosed by these markers and compare it with the ground-truth count. If such parsing fails, we fall back to checking whether the ground-truth count appears anywhere in the model output, either as an Arabic numeral or as an English number word. This heuristic allows us to robustly score free-form responses while preserving the intended answer format. An example QA pair is shown below.

% \begin{quote}
% \textbf{Question:} How many people are there in the image? Response Example: There are \texttt{**<number>**} of people in the image. \\
% \textbf{Answer:} There are \texttt{**5**} of people in the image.
% \end{quote}
\subsubsection{Understanding evaluation.}
We evaluate counting understanding using 540 QA pairs constructed from 
the test split of PixMo-Count. Each prompt includes a \texttt{Response Example} 
field that instructs the model to place the predicted count inside markdown 
markers \texttt{** **}. Each QA pair follows the template:
\begin{quote}
\textbf{Question:} \texttt{"How many \{OBJECT\}s are there in the image? Response Example: There are **<COUNT>** of \{OBJECT\}s in the image."} \\
\textbf{Answer:} \texttt{"There are **\{COUNT\}** of \{OBJECT\}s in the image."}
\end{quote}
\noindent During evaluation, we parse the substring enclosed by the markers 
and compare it with the ground-truth count. If parsing fails, we fall back 
to checking whether the ground-truth count appears anywhere in the model 
output, either as an Arabic numeral or as an English number word.

\subsubsection{Generation evaluation.} 
Following the prompt generation protocol of GenEval~\cite{ghosh2023geneval}, 
we construct a total of 90 prompts by combining 10 object classes 
and 9 count words, yielding 90 prompts in total. 
Each prompt follows the template:

\begin{quote}
\texttt{"a photo of \{COUNT\} \{OBJECT\}s"}
\end{quote}

\noindent where \texttt{OBJECT} $\in$ \{\texttt{person, dog, cat, car, 
cup, chair, book, bottle, apple, tie}\} and \texttt{COUNT} $\in$ 
\{\texttt{two, three, four, five, six, seven, eight, nine, ten}\}.

For each prompt, we generate four images with different random seeds 
and report the average accuracy across all generated samples.
To obtain object counts from generated images, we follow same detection-based approach as GenEval~\cite{ghosh2023geneval}.

\subsection{Spatial Relation}
\subsubsection{Dataset detail.}
Our \textit{visual instruction tuning dataset} for spatial relation understanding consists of 200K image-question-answer pairs. We construct composite images using object crops drawn from the ImageNet-1K~\cite{deng2009imagenet} training set. Specifically, for each sample, we randomly select two images from different ImageNet classes and paste them onto a $512 \times 512$ canvas with a random color background. We resize each selected image to a random size~(between 150 and 300 pixels per side, with an aspect ratio constrained near 1:1) and place them such that the two images do not overlap and exhibit a clearly non-ambiguous relative spatial relationship. 

We consider four relative positions—\texttt{top-left}, \texttt{top-right}, \texttt{bottom-left}, and \texttt{bottom-right}—defined by the difference between the centroids of the two images, and enforce an even distribution of 50K samples per position. For each sample, a question-answer pair is generated by randomly selecting one of 10 diverse QA templates that ask about the relative position of object $A$ with respect to object $B$. Each question is appended with a \texttt{Response format} hint that specifies the expected answer structure. For the \textit{spatial-relation-aware generation dataset}, we construct the training set from the same images used in the understanding split. We leverage the spatial relation annotations from the understanding training set and use Qwen3-VL~\cite{bai2025qwen3} to generate spatially-aware generation prompts conditioned on these annotations. Table~\ref{tab:spatial_example} presents representative samples from both the spatial relation understanding and generation training sets.

Each QA pair follows the template: 
\begin{quote}
\textbf{Question:} \texttt{"\{QUESTION\}\textbackslash nResponse format: \{RESPONSE\_FORMAT\}"} \\ 
\textbf{Answer:} \texttt{"\{ANSWER\}"} 
\end{quote} 

\noindent For example, given a template sampled at random: \begin{quote} \textbf{Question:} \texttt{"Where is the \{OBJECT\_A\} located relative \\ to the \{OBJECT\_B\}?} \\ \texttt{Response format: The \{OBJECT\_A\} is located in the \\ top-left/top-right/bottom-left/bottom-right of the \{OBJECT\_B\}."} \\ \textbf{Answer:} \texttt{"The \{OBJECT\_A\} is located in the \{POSITION\} of the \{OBJECT\_B\}."} \end{quote} 

\noindent where \texttt{POSITION} $\in$ \{\texttt{top-left, top-right, bottom-left, bottom-right}\}, and \texttt{OBJECT\_A}, \texttt{OBJECT\_B} are ImageNet class names.
\input{x_supple/tabs/dataset_spatial}
\subsubsection{Understanding evaluation.} 
\subsubsection{Generation evaluation.} 
\input{x_supple/figs/relation_prompts}
We generate 100 prompts describing spatial relationships between two objects, following the same answer-style declarative format used during training. Each prompt is a sentence that states the diagonal relation between two object classes, \textit{e.g.}, ``The truck can be found to the top-left of the baseball.'' The full list of evaluation prompts is provided in Fig.~\ref{fig:spatial_prompts}. For each prompt, we generate four images with different random seeds and report the average accuracy across all generated samples.

For evaluation, we follow the same detection-based approach as GenEval~\cite{ghosh2023geneval}, obtaining bounding boxes for both objects. 
The relative position between objects is determined by the difference 
between bounding box centroids: if object $A$ is centered at $(x_A, y_A)$ 
and object $B$ is centered at $(x_B, y_B)$, following standard image 
coordinate conventions where the $y$-axis increases downward, then:
\begin{itemize}[label=\textbullet]
    \item \textbf{top-left:} $x_A < x_B$ and $y_A < y_B$
    \item \textbf{top-right:} $x_A > x_B$ and $y_A < y_B$
    \item \textbf{bottom-left:} $x_A < x_B$ and $y_A > y_B$
    \item \textbf{bottom-right:} $x_A > x_B$ and $y_A > y_B$
\end{itemize}
A sample is counted as correct if the detected relation matches the one 
specified in the prompt. If the detector fails to localize either object, 
the sample is excluded from evaluation.


\subsection{Text Recognition and Generation}
\subsubsection{Dataset detail.} We construct the synthetic OCR training set from a public English sentence corpus on Hugging Face. Specifically, we use the training split of \path{agentlans/high-quality-english-sentences} and keep the first 200K samples. Each sample is normalized by replacing line breaks with spaces, trimming whitespace, and truncating the target string to at most 120 characters. The normalized text is then formatted as \texttt{\# <sentence>}, \textit{i.e.}, a single level-1 markdown heading rather than a full free-form document. We convert this markdown string to HTML using the Python \texttt{markdown} library with standard extensions for fenced code blocks, tables, automatic line breaks, and list handling, and then rasterize the HTML into a PNG image with \texttt{wkhtmltoimage}. We define four hand-designed CSS templates, \texttt{clean\_light}, \texttt{dark}, \texttt{sepia}, and \texttt{blue\_gray}, which vary the background, text color, and font family. During understanding training, one of these templates is sampled at random for each example, whereas generation training uses the \texttt{clean\_light} template throughout. Figure~\ref{fig:ocr_dataset} shows representative examples.

We use the same underlying text-image pair for both modalities. For understanding, each rendered image is paired with the fixed instruction \texttt{Extract all text from the image.}, and the normalized sentence is used as the target. For generation, the same sentence is embedded into a fixed caption template describing a Mathpix-style rendering with sharp, legible text in natural reading order. This construction lets us compare \textit{und} and \textit{gen} transfer on identical textual content while changing only the direction of supervision. The explicit understanding and generation formats are shown below.

\begin{quote}
\textbf{Understanding question:} Extract all text from the image. \\
\textbf{Understanding answer:} The train arrived before sunset. \\
\textbf{Generation prompt:} A Mathpix Markdown format with sharp, legible black text. High-resolution typography, top-down view. \\
The text is rendered in natural left-to-right, top-to-bottom reading order. The text reads: The train arrived before sunset.
\end{quote}

\subsubsection{Understanding Evaluation.} We evaluate text recognition on the first 250 samples from the test split of the same corpus. For each example, we normalize the reference sentence in the same way as during training and re-render it as a \texttt{clean\_light} markdown image. At evaluation time, we use the explicit instruction \texttt{Extract all text from the image in reading order.} and otherwise follow the default understanding prompt format and decoding procedure of each model. After generation, both prediction and reference are lowercased, line breaks are replaced with spaces, and leading and trailing whitespace is stripped. We report WER, CER, BLEU, METEOR, average edit distance, and token-level precision, recall, and F1.

\subsubsection{Generation Evaluation.} We evaluate text generation on the first 250 samples from the test split. For each target sentence, we construct the same Mathpix-style caption template shown above and generate an image following the default text-to-image inference setting of each model. We then apply GLM-OCR~\cite{glm_ocr} to the generated image and extract text. The extracted text and the ground-truth sentence are normalized with the same lowercasing and whitespace processing used in understanding evaluation, and we report the same set of metrics: WER, CER, BLEU, METEOR, average edit distance, and token-level precision, recall, and F1.

\clearpage
\section{Additional Quantitative Results}
In this section, we provide the complete quantitative results for all three tasks under both objectives: understanding and generation.
\label{sec:appendix_additional_quantitative}
\subsection{(Counting)} 
\input{table/appendix_counting_full}
Table~\ref{tab:appendix_counting_full} extends the counting transferability results in Table~\ref{tab:transfer} of the 
main paper by additionally reporting within-modality training results (\textit{und}$\rightarrow$\textit{und} and \textit{gen}$\rightarrow$\textit{gen}), 
which serve as upper bounds for the corresponding transfer directions.
For generation, \textit{und}$\rightarrow$\textit{gen} transfer proves surprisingly competitive against the 
within-modality upper bound. In Lumina-DiMOO, \textit{und}$\rightarrow$\textit{gen} transfer achieves \textit{gen}$\rightarrow$\textit{gen}, and in BLIP3-o, it substantially outperforms the within-modality baseline.
For understanding, \textit{gen}$\rightarrow$\textit{und} transfer is consistently observed across architectures 
but falls short of the within-modality upper bound (\textit{und}$\rightarrow$\textit{und}) 
in all cases, indicating that transferability exists in this direction while leaving a gap 
relative to direct understanding supervision.

\subsection{Spatial Relation}
\subsection{Text Recognition and Generation}
\input{table/appendix_spatial_ocr_full_combined}

\clearpage
\section{Generation Training Results Across Unified Multimodal Models}

\label{sec:appendix_gen_results}
\vspace{-20pt}
\input{table/appendix_fid_is_combined}
\vspace{-20pt}

We measure IS and FID under two prompt settings to assess how generation-side fine-tuning affects the output distribution of the baseline generator. First, we evaluate with task-specific prompts that match the format used during generation training. We use 1.8K samples for counting, 2K for spatial relation, and 1.8K for text recognition and generation. Second, to examine the effect on more general image synthesis, we repeat the evaluation using 1K prompts from the COCO validation set. The results are reported in Table~\ref{tab:appendix_fid_is_combined}.
We first consider FID. Across both prompt settings and all tasks, \textit{und}$\rightarrow$\textit{gen} consistently yields lower FID than \textit{gen}$\rightarrow$\textit{gen}. Because FID is computed with respect to the baseline model's generations under the same prompt set, this result indicates that understanding-based tuning preserves the baseline generator's original output distribution more faithfully than direct generation tuning.

By contrast, IS is considerably more task-dependent. Under the task-specific prompt setting, the IS of \textit{und}$\rightarrow$\textit{gen} is generally close to that of the baseline, whereas \textit{gen}$\rightarrow$\textit{gen} often attains higher IS on counting. One possible explanation is that direct generation tuning on the counting dataset, which contains multiple instances of the same object category, makes object appearance more classifier-friendly within the narrow counting prompt distribution. However, this tendency does not extend consistently to spatial relation or text recognition and generation, and under the COCO validation prompt set the apparent advantage of \textit{gen}$\rightarrow$\textit{gen} weakens or disappears. We therefore interpret changes in IS not as direct evidence of improved general image quality, but rather as reflecting adaptation to a narrow task-specific prompt distribution. 

Overall, these results suggest that \textit{und}$\rightarrow$\textit{gen} perturbs the baseline image-generation distribution substantially less than \textit{gen}$\rightarrow$\textit{gen}.

\clearpage
\section{Additional Qualitative Results}
\label{sec:appendix_qualitative}
\subsection{Counting}
\subsection{Spatial Relation}
\subsection{Text Generation}
