\section{Introduction}

% Multimodal Large Language Models (MLLMs) and diffusion-based generative models~\cite{} have each achieved remarkable progress in recent years. Despite their respective successes, these two model families have evolved largely in isolation: one focuses on interpreting the visual world, and the other on generating it. This naturally motivates the constrcution of a unified model capable of jointly performing understanding and generation. In response, recent Recent Unified Multimodal Models (UMMs)~\cite{} have emerged as a early attempts to integrate these capabilities within a single architecture, achieving performance comparable to expert models in each domain. However, this progress leads to a fundamental inquiry: is the goal of UMMs simply to consolidate two expert systems into one parameter space, or to achieve genuine synergy between understanding and generation that enables a deeper form of world modeling?

% % To address this question, we must first determine whether meaningful synergy actually emerges within UMMs. 
% Several prior works suggest that complementarity between perception and generation is possible. For example, REPA improves diffusion generation by aligning intermediate DiT representations with DINOv2 features, and other works have shown that auxiliary reconstruction objectives can enhance visual understanding in MLLMs. However, it remains unclear whether such synergy naturally arises within unified architectures. While some studies report positive interactions between tasks, others observe interference, where joint modeling degrades performance. We argue that this discrepancy stems from how synergy has been evaluated. Existing approaches typically rely on aggregate performance changes on general multimodal benchmarks, which do not directly measure whether knowledge learned in one task influences the representations used for another. True synergy cannot be reduced to parameter sharing; it requires knowledge integration, where knowledge acquired from one task meaningfully shapes the representations underlying the other. Therefore, evaluating synergy in UMMs demands a more direct measurement of cross-task knowledge transfer.

% To this end, we propose task transferability as a concrete indicator of knowledge integration. Specifically, we examine whether learning in one modality, understanding or generation, can directly improve performance in the other without explicit supervision in that target modality.  As a testbed, we focus on counting, a capability known to be challenging for both MLLMs and diffusion models. We measure whether counting-focused visual instruction tuning improves counting-aware image generation, and conversely, whether counting-aware generative training enhances counting performance in understanding. Through systematic experiments across representative UMM architectures, we find a clear architectural distinction between models that exhibit bi-directional transferability and those that do not. Transferability consistently emerges only in architectures equipped with a shared transformer backbone and a unified image tokenizer. In contrast, loosely coupled designs fail to demonstrate meaningful cross-modal improvement. These findings suggest that knowledge integration is not an automatic consequence of joint training, but is fundamentally shaped by architectural design, highlighting the importance of structural coupling for enabling synergy.

% Finally, we extend our analysis beyond counting to investigate how transferability manifests across tasks of varying granularity. From fine-grained tasks such as OCR to coarse semantic reasoning such as spatial position estimation, we observe that different tasks exhibit modality preference—some are more effectively learned through understanding and then transferred to generation, while others benefit from the reverse direction. Notably, in several cases, training in the modality more naturally aligned with a task and subsequently transferring to the target modality yields greater efficiency than direct supervision in the target itself. These findings reveal that the relationship between understanding and generation is neither trivial nor symmetric, but depends on both task characteristics and architectural structure. Our study thus reframes UMM design not as simple functional unification, but as the deliberate construction of architectures capable of genuine knowledge integration
Multimodal understanding and image generation have each made remarkable progress in recent years, yet they have mostly evolved in isolation. This separation has recently motivated the development of Unified Multimodal Models (UMMs), which aim to integrate both capabilities within a single architecture. Recent UMMs demonstrate that such unification can achieve performance comparable to specialized systems for both tasks \cite{XXX}.
However, despite this progress, it remains unclear how the two capabilities interact once they are trained within a single model. In particular, the potential benefits that arise from jointly modeling understanding and generation remain underexplored.

% However, an important question remains unresolved: do these models truly integrate understanding and generation, or do they simply host two loosely connected systems within shared parameters? 

% Recent UMMs achieve performance comparable to dedicated expert models in both tasks, suggesting that unification incurs little compromise. Yet a fundamental question remains: "do these models truly unify the two capabilities, or do they merely host two independent systems within shared parameters?'' If understanding and generation are genuinely integrated, knowledge acquired through one task should be accessible to the other. However, whether such cross-task knowledge sharing actually occurs in current UMMs has not been directly examined.

% For instance, aligning generative representations with discriminative visual features has been shown to improve image synthesis \cite{XXX}, while auxiliary reconstruction objectives can enhance visual understanding in multimodal language models \cite{XXX}. These findings suggest that the two objectives may influence one another through shared representations. However, existing analyses typically evaluate such interactions through aggregate performance changes on general benchmarks. Improvements observed under joint training may arise from factors such as additional supervision or regularization effects, rather than from genuine knowledge sharing between tasks. As a result, it remains difficult to determine whether knowledge learned for one task is actually utilized by the other.

Several recent studies have begun to investigate interactions between understanding and generation in multimodal systems. For example, some works demonstrate that semantic representations from understanding can improve the semantic fidelity of generation, while fine-grained visual representations learned through generation may enhance grounding and compositional reasoning \cite{wu2026liquid,xin2025lumina,yang2025mmar,wu2025harmonizing,kou2025orthus,li2025dual,zhang2025unified}.  Despite these insights, existing analyses typically examine interactions between understanding and generation through aggregate performance changes on general benchmarks. While these results provide useful high-level observations, they do not directly reveal how knowledge is shared between the two capabilities. Furthermore, because such evaluations rely on broad benchmark improvements, it remains difficult to determine whether the observed gains stem from genuine cross-task knowledge transfer or from indirect factors such as additional supervision or regularization effects. 

We address this gap by studying a more direct and actionable signal of cross-task interaction: \emph{transferability}.
Instead of inferring interaction from aggregate benchmark gains, we ask whether training a given task (\textit{e.g.,} counting) in one modality (understanding or generation) reliably improves the other modality on the same task, under controlled interventions.
This perspective enables (i) diagnosing architectural conditions under which understanding--generation coupling produces meaningful knowledge sharing, and (ii) leveraging such transfer as a practical mechanism to inject target capabilities for the given tasks from understanding into generation.
% without corrupting the pretrained generative distribution.

% Several prior works have explored interactions between understanding and generation in multimodal models. However, these studies typically measure synergy through aggregate performance changes on general benchmarks, which do not directly reveal whether knowledge is actually shared across tasks. A performance gain from joint training could stem from regularization effects or increased data exposure rather than genuine knowledge transfer. To address this, we propose a more direct measurement: we train a model on a specific task in one modality alone and test whether the other modality improves on the same task without any explicit supervision. We term this property \textit{transferability}, and adopt it as a concrete indicator of cross-task knowledge integration in UMMs. 

Specifically, we begin with a controlled analysis using \textit{counting} as a probe task. Across several representative Unified Multimodal Models, we observe that transferability strongly depends on architectural design. In particular, models that employ a fully shared transformer backbone with a unified visual tokenizer exhibit consistent transfer between modalities, whereas architectures that leverage a separated backbone or a tokenizer show little or no transfer. These results suggest that the degree of representation sharing plays a critical role in enabling cross-task knowledge transfer.

Building on this observation, we investigate whether transferability can be used as a practical mechanism for improving generative capabilities. Directly training generation to acquire a specific capability—such as accurately counting objects—can be challenging, as generation must simultaneously preserve image quality and the underlying data distribution. Instead, we find that training the corresponding \textbf{understanding task} can effectively introduce the capability into generation through transfer. This strategy allows generation performance to improve on the target capability while largely preserving generative quality.

Finally, we evaluate this phenomenon across multiple visual capabilities, including counting, spatial relation reasoning, and OCR. Our results show that cross-modal transfer can consistently improve generation performance on these tasks while maintaining overall multimodal understanding ability. At the same time, we observe that the degree of transfer varies across tasks, highlighting differences in how visual capabilities are represented across modalities.


We begin by investigating whether transferability exists in current open-weight UMMs. To this end, we categorize existing architectures into four representative paradigms and select a model from each. Using counting as a controlled visual concept, we train understanding with counting visual question–answer data and generation with counting-aware text–image pairs, and then measure how well numerical concepts transfer to the untrained modality. Through these experiments, we identify architectures where transferability is inherent and others where it is absent. In particular, models in which understanding and generation share a unified visual encoder within a shared transformer exhibit strong bi-directional transferability, whereas architectures lacking this strong structural coupling show limited transferability.

We then turn to a natural follow-up question: when transferability emerges, is it consistently bi-directional across tasks? Prior works suggests that understanding objectives tend to promote high-level semantic representations, while generation objectives often capture fine-grained visual structures. This distinction raises the possibility that transferability may not behave uniformly across tasks. To examine this, we select representative tasks with distinct visual characteristics and evaluate transfer in both directions. Our experiments show that transfer is neither uniform nor symmetric across tasks. We find that tasks relying on coarse, image-level semantics exhibit strong bi-directional transferability, whereas tasks requiring fine-grained visual discrimination show more limited and asymmetric transfer.


Our contributions ar as follows:
\begin{itemize}
    \item first
    \item second
    \item third
\end{itemize}