\section{Related Work}
\subsection{Image Understanding and Generation in UMMs}
% 모델 아키텍쳐 위주로
The success of large language models~\cite{floridi2020gpt, touvron2023llama} has catalyzed the extension of language modeling into the multimodal domain. Multimodal large language models (MLLMs)~\cite{bai2025qwen3,  } enable strong image understanding by pairing a visual encoder with an LLM. More recently, Unified Multimodal Models (UMMs) have pushed this paradigm further by incorporating image generation alongside understanding within a single model~\cite{team2024chameleon, wu2025janus, deng2025bagel, xin2025lumina, chen2505blip3}.

\subsubsection{Visual representations.}
A key design axis in UMMs is the choice of visual representation for encoding input images. One line of work adopts language-supervised encoders such as CLIP~\cite{radford2021clip} and SigLIP~\cite{zhai2023siglip}, whose representations are aligned with text and thus preserve high-level semantics amenable to reasoning. A complementary line leverages autoencoders originally designed for generation, such as VQ-VAE~\cite{vandenoord2017vqvae} or continuous VAE~\cite{kingma2022vae}, which prioritize pixel-level reconstruction fidelity. Analyses in Janus~\cite{chen2025januspro,wu2025janus} and related works~\cite{cambrian} show that language-supervised representations substantially outperform compression-oriented tokenizers for understanding tasks, motivating several UMMs to employ separate encoders for each modality direction.

\subsubsection{Generation objectives.}
Image generation in UMMs follows two main paradigms. \emph{Autoregressive} models~\cite{team2024chameleon, yu2024emu3} employ discrete image tokenizers and generate tokens sequentially, directly extending the next-token prediction framework of LLMs. \emph{Diffusion}-based models instead learn a denoising objective~\cite{ho2020ddpm, song2020ddim, lipman2023flowmatching} and can be further divided into discrete-token diffusion~\cite{yang2025mmada, xin2025lumina}, which applies masked or absorbing-state diffusion over quantized codes, and continuous diffusion~\cite{zhou2024transfusion, chen2505blip3}, which operates on continuous latent representations.

\subsubsection{Architectural taxonomy.}
We identify three dominant architecture families (Figure~\ref{fig:arch_categorize}). \textbf{(a) Single-transformer} designs route all modalities through a shared transformer backbone, maximizing parameter sharing~\cite{team2024chameleon, yu2024emu3, yu2025emu35}. \textbf{(b) Decoupled-transformer} designs employ separate transformers for language and vision, where the LLM provides conditioning signals (\eg, hidden states) that are subsequently fed into a dedicated diffusion module for generation~\cite{dong2025metaquery, tian2025metamorph}. \textbf{(c) Mixture-of-Transformers (MoT)} designs maintain modality-specific transformer branches but couple them through joint or cross-attention layers, allowing selective parameter sharing while retaining specialized capacity~\cite{deng2025bagel, li2025unifork, wang2025lightfusion, fan2025hbridge}. As we show in Section~\ref{sec:umms}, this architectural choice has a decisive impact on whether knowledge transfers across understanding and generation. We visually present the architectural trend in Figure.~\ref{fig:arch_categorize}


\subsection{Understanding Interactions Between Understanding and Generation in UMMs}
% ablation study / benchmark 위주의 분석
A central question in designing Unified Multimodal Models is whether multimodal understanding and generation genuinely benefit each other within a shared architecture. Several works report potential synergy, arguing that high-level semantic representations learned for understanding can improve the semantic fidelity of generation, while fine-grained visual representations acquired through generation may enhance visual grounding and compositional reasoning~\cite{wu2026liquid, xin2025lumina, yang2025mmar, wu2025harmonizing, kou2025orthus, li2025dual,zhang2025unified}. On the other hand, research such as UniFork~\cite{li2025unifork} empirically demonstrates that ideal representations for these two tasks throughout the network layers are distinct, suggesting that simple parameter sharing can lead to a representational compromise. This perspective aligns with a broader trend of decoupled architectures that utilize task-specific branches in higher layers or separated visual paths to mitigate potential conflicts~\cite{wu2025janus, chen2025januspro, deng2025bagel}. 

While these studies provide valuable insights into the potential benefits and trade-offs of joint training, most analyses examine interaction through aggregate performance changes on general benchmarks. For example, prior work often evaluates whether increasing generation-oriented training improves overall understanding benchmarks (e.g., MMMU or POPE), or conversely whether additional understanding supervision improves generation benchmarks such as GenEval. Although such evaluations provide useful high-level signals, they make it difficult to isolate how specific capabilities learned in one modality influence the other. In contrast, our work investigates cross-task interaction at the capability level through the lens of transferability. Rather than inferring interaction from aggregate benchmark improvements, we directly examine whether learning a capability in one modality improves the same capability in the other modality. 