\section{Does knowledge transfer emerge in UMMs?}
\input{fig/architecture}
\label{sec:umms}

% We now move from qualitative observations to a task-level analysis of cross-modal transfer. We design experiments around quantity-aware reasoning. On the understanding side, the task is object \textbf{counting}. On the generation side, the task is \textbf{count-conditioned image synthesis}. Both tasks share the same underlying concept of object quantity, providing a common evaluation axis for measuring whether numerical knowledge learned from one objective transfers to the other. We set up two complementary training conditions: training with only the understanding objective and evaluating generation (Und.\ $\rightarrow$ Gen.), and vice versa (Gen.\ $\rightarrow$ Und.).



In this section, we investigate whether knowledge transfer between understanding and generation emerges in current open-source UMMs. Our goal is to examine whether learning a capability in one modality~(understanding or generation) can improve the same capability in the other modality under controlled interventions, and whether the UMM architecture is advantageous for the transfer. To this end, we conduct a task-level analysis using counting as a probe task, which provides a clean and quantifiable concept that both understanding and generation must encode using one representative open-source UMM from different architectural categories of UMMs.

% In this section, we examine whether knowledge transfer between understanding and generation emerges in current open-source UMMs. Following the categorization in Figure.~\ref{fig:arch_categorize}, we select a representative open-source model from each paradigm: BAGEL~\cite{deng2025bagel} for (a), Janus-Pro~\cite{chen2025januspro} for (b), Lumina-DiMOO~\cite{xin2025lumina} for (c), and BLIP3-o~\cite{chen2505blip3} for (d).
% We leverage counting as the probe task since it offers a clean, quantifiable concept that both modalities must encode. The model must be capable of counting the number of objects present, regardless of whether it is answering a question or generating an image. For each model, we conduct two separate training runs using LoRA-based supervised fine-tuning. In the understanding-only condition, the model is trained on counting visual question answering data. In the generation-only condition, it is trained on counting-aware text-to-image pairs constructed from the same source images.

\vspace{-10pt}
\subsection{Experimental setup} 

% Following the architectural categorization in Figure~\ref{fig:arch_categorize}, we select a representative open-source model from each category: BAGEL~\cite{deng2025bagel} for (a), Janus-Pro~\cite{chen2025januspro} for (b), Lumina-DiMOO~\cite{xin2025lumina} for (c), and BLIP3-o~\cite{chen2505blip3} for (d). 
Considering the diverse architectural choices introduced in Section~\ref{subsec:taxonomy}, we select a representative open-source model: BAGEL~\cite{deng2025bagel} for (\textbf{a) Mixture of Transformers}, Janus-Pro~\cite{chen2025januspro} for (\textbf{b) Shared Transformer with separated image encoder}, Lumina-DiMOO~\cite{xin2025lumina} for (\textbf{c) Shared Transformer with unified image encoder}, and BLIP3-o~\cite{chen2505blip3} for \textbf{(d) Separated Transformer}. The architectural overview of each selected model is illustrated in Figure~\ref{fig:arch_categorize}. For each model, we conduct two separate training runs using LoRA-based supervised fine-tuning. When learning understanding, the model is trained on counting visual question answering data. When learning generation, the model is trained on counting-aware text-to-image pairs constructed from the same source images.

Specifically, for both understanding and generation, we construct training sets based on PixMo-Points~\cite{deitke2025molmo}. For understanding, we filter PixMo-Points, originally designed for point coordinate prediction, to retain images with object counts ranging from 0 to 20 and reformulate them into counting-based visual question answering data. For generation, we use Qwen3-VL~\cite{bai2025qwen3} to produce image descriptions for the same images and append object count information to form generation prompts.
\input{table/counting_models}
\input{fig/counting_qual}
For evaluation, we use two separate protocols. Understanding is evaluated on the test split of PixMo-Count~\cite{deitke2025molmo} in a VQA format, where the model is asked to count target objects in an image and responds with a number. For generation, we follow the counting protocol from GenEval~\cite{ghosh2023geneval} and construct prompts in the form \texttt{"A photo of [NUMBER] [OBJECT]s"} with counts ranging from 2 to 10. Generated images are evaluated using an object detector~\cite{cheng2022masked} to verify whether the number of detected objects matches the specified count. We report both accuracy and mean absolute deviation (MAD) to capture exact correctness and proximity to the ground truth.


\vspace{-10pt}
\subsection{\textbf{Transferability across architectures}}
% \subsubsection{Bi-directional transfer emerges in models with shared representations.} 
\subsubsection{Transferability varies across architectures.}
Table~\ref{tab:transfer} summarizes the results. The most consistent finding is that knowledge transfer from understanding tasks to generation, which we term as \textit{und-to-gen transfer}~(\textit{und} $\rightarrow$ \textit{gen}) is observed in all models except Janus-Pro~\cite{chen2025januspro}. BAGEL~\cite{deng2025bagel}, Lumina-DiMOO~\cite{xin2025lumina}, and \mbox{BLIP3-o~\cite{chen2505blip3}} all show improved generation accuracy and lower MAD when trained solely with the understanding objective, indicating that numerical knowledge acquired solely through understanding can be transferred to generation. On the other hand, we observe that the inverse direction, \mbox{\textit{gen-to-und transfer}~(\textit{gen} $\rightarrow$ \textit{und}),} is more selective: only BAGEL and Lumina-DiMOO exhibit bi-directional transfer, where generation-only training also yields gains in counting understanding. Among all models, Lumina-DiMOO shows the strongest effect in both directions, underscoring that transferability is not uniform across UMM architectures.  Qualitative examples illustrating this behavior are shown in Figure~\ref{fig:counting_quals}.
% Among all models, Lumina-DiMOO shows the strongest effect in both directions. Interestingly, training Lumina-DiMOO with only the understanding objective yields even greater improvement in generation than direct generation supervision~\mbox{(\textit{gen}$\rightarrow$
% \textit{gen}),} underscoring that transferability is not uniform across UMM architectures and that, for certain designs, indirect supervision through a related task can be more effective than task-specific training.

\vspace{-10pt}
\subsubsection{Which architecture results in the strongest transferability?} From an architectural perspective, we observe that models which employ a fully shared transformer backbone for both understanding and generation, along with a unified vision encoder (category (c) in Figure~\ref{fig:arch_categorize}), such as Lumina-DiMOO~\cite{xin2025lumina}, exhibit the strongest transferability. This result is consistent with recent empirical findings and further explains why adopting a unified encoder for image understanding and generation produces synergistic effects in UMMs~\cite{wu2025harmonizing, jiao2025unitoken, wu2026liquid}: modeling both tasks within a single transformer backbone reduces the representational discrepancy between the two objectives and promotes better alignment. When both understanding and generation operate over a shared representation, knowledge acquired through one objective naturally propagates to the other, facilitating cross-task transfer. By contrast, architectures that maintain separate visual pathways for each task offer fewer opportunities for such propagation, which explains the weaker or absent transfer observed in models like Janus-Pro.