
\section{\textbf{Using Transferability to Improve Generative Capabilities}
}
In this section, we investigate how the transferability emerged in UMMs can be practically exploited. In practice, generative models often exhibit weaknesses in specific capabilities such as spatial reasoning or object counting. A straightforward approach to address these limitations is to perform additional fine-tuning on generation data targeting the desired capability. However, improving a specific skill while preserving overall image generation quality is often difficult in practice: direct generation fine-tuning can easily degrade generative metrics or distort visual quality.

We propose a practical alternative by leveraging transferability. Instead of directly optimizing the generation objective, we introduce the target capability through the understanding modality and allow the learned knowledge to transfer into generation. This approach enables targeted improvements in generative capabilities while maintaining the overall generation quality of the model. To study this mechanism, we analyze understanding-to-generation transfer across three representative tasks, \textit{i.e.,} counting, spatial relation reasoning, and OCR, through the following sections.


% The experiments in Section~\ref{sec:umms} established that knowledge transfer between understanding and generation occurs in UMMs with certain architectural properties, and that for counting, this transfer is bi-directional. A natural follow-up question is whether similar patterns hold for other tasks. We investigate this by extending our analysis to two additional domains: OCR, where understanding and generation share character-level visual patterns, and spatial reasoning, where both tasks rely on relative object locations. Throughout this section, we use Lumina-DiMOO~\cite{xin2025lumina} with LoRA-based supervised fine-tuning, as it exhibited the strongest cross-task transfer in the preceding experiments.

\subsection{Counting}
% \input{table/counting_undgen}
\input{table/gen_fid}
\input{fig/undgen_counting}

We first revisit the counting generation from this perspective. Table~\ref{tab:transfer} previously showed that counting exhibits strong transferability from understanding to generation. To examine whether this transfer can be practically beneficial, we compare the model trained through understanding (\textit{und} $\rightarrow$ \textit{gen}) with a model directly trained on counting generation data (\textit{gen} $\rightarrow$ \textit{gen}) in Table~\ref{tab:gen_quality}. Surprisingly, the model trained through understanding achieves a lower mean absolute deviation (MAD) in counting generation, indicating that it produces images with object counts closer to the target number. In other words, introducing counting knowledge through the understanding modality leads to more accurate counting behavior in generation than directly optimizing the generation objective itself. 

The qualitative results in Figure~\ref{fig:undgen_counting} further highlight this difference. When the model is trained on counting-aware text–image pairs (\textit{gen} $\rightarrow$ \textit{gen}), the narrow distribution of the fine-tuning data causes the model to drift from its original training distribution, resulting in noticeable degradation of overall image quality. In contrast, the model trained via understanding (\textit{und} $\rightarrow$ \textit{gen}) preserves the overall image quality while selectively improving the numerical concept required for counting. To quantify this observation, we additionally measure image quality using FID and IS in Table~\ref{tab:gen_quality}. Training directly on generation data (\textit{gen} $\rightarrow$ \textit{gen}) degrades both metrics compared to the baseline model, confirming the loss of generative quality. In contrast, the understanding-based approach maintains generation quality while improving counting accuracy, demonstrating that transfer through understanding provides a more stable way to introduce new capabilities into generation.

\subsection{Spatial relation}
\label{sec:spatial_rel}
We next consider a structurally different capability, spatial reasoning, where the model must encode the relative layout of objects. Unlike counting, this task requires capturing compositional relationships between multiple entities. This allows us to examine whether the transferability observed in counting extends to more complex relational concepts.

\input{table/position_lumina}
\input{fig/spatial_undgen}

\subsubsection{Experimental setting.} We observe that pre-trained models already handle simple spatial relations such as left, right, top, and bottom reasonably well. We therefore focus on a more challenging setting: diagonal relations between two objects (e.g., top-left, top-right, bottom-right, bottom-left). Since existing spatial reasoning datasets~\cite{} lack the scale and complexity needed for fine-tuning, we construct a synthetic dataset of 200K samples from ImageNet~\cite{deng2009imagenet}. For each sample, as shown in Figure~\ref{fig:spatial_dataset} we select two class–image pairs and place them on a canvas according to a randomly sampled diagonal direction. Each sample is paired with a prompt of the form \texttt{"[class A] is located in the [direction] of [class B]"}.
\input{fig/spatial_dataset}

For evaluation, generation is evaluated by constructing prompts in the same format as the training data and, following GenEval~\cite{ghosh2023geneval}, applying a detection model~\cite{cheng2022masked} to localize bounding boxes of the two objects in the generated image. We compute the center points of the detected boxes and verify whether their relative position matches the specified direction. More details of the dataset construction and evaluation pipeline can be found in Section~\ref{?}

\subsubsection{Results.}
Table~\ref{tab:gen_quality_spatial} confirms that understanding-to-generation transferability also holds for spatial relation. The model trained through understanding (\textit{und} $\rightarrow$ \textit{gen}) achieves higher spatial accuracy than the baseline, demonstrating that relational knowledge acquired through understanding transfers to generation. Training directly on generation data (\textit{gen} $\rightarrow$ \textit{gen}) yields slightly higher accuracy, but at the cost of degraded FID and IS scores. The qualitative results in Figure~\ref{fig:spatial_pos} illustrate this trade-off more clearly. (\textit{gen} $\rightarrow$ \textit{gen}) easily shifts to the training data distribution, producing images that resemble cropped objects pasted onto a canvas rather than natural scenes. In contrast, the understanding-trained model improves spatial correctness while preserving the realistic appearance of the generated images.

% Figure~\ref{fig:spatial_pos} summarizes the results. Training with the understanding objective alone improves generation accuracy. Training with the generation objective alone similarly improves understanding. The magnitude of transfer is comparable in both directions. This pattern is consistent with the quantity-aware reasoning results in Section~\ref{sec:umms} and contrasts with the asymmetric transfer observed in text recognition and generation.

\subsection{Text recognition and generation}
We first examine text recognition~(OCR) and text-contained image generation as a pair of complementary tasks. Both tasks operate on character-level visual cues such as strokes, spacing, and glyph structure, requiring the model to capture highly localized and fine-grained visual details. We investigate whether learning such details through one objective transfers to the other.

\input{fig/ocr_dataset}
\input{fig/ocr_undgen}
\subsubsection{Experimental setting.} We construct a dataset of 200K image-text pairs by rendering markdown documents into images. Existing OCR benchmarks~\cite{} tend to focus on either short sign-level text in natural scenes or dense full-page documents. Neither setting allows meaningful variation in both recognition and generation. Our rendered dataset provides diverse text content at a moderate difficulty level, where both tasks can exhibit measurable room for improvement. 
For evaluation, we report five metrics. word error rate (WER) and character error rate (CER) measure exact correctness at the word and character level. Edit Distance captures the minimum number of operations needed to align the prediction with the ground truth. METEOR~\cite{banerjee2005meteor} and F1 assess partial matches by accounting for token overlap and ordering. For understanding, these metrics are computed between the model's output and the ground-truth text. For generation, we first apply an OCR model~\cite{glmocr} to read text from the generated image and then compute the same metrics against the original prompt. More details of the dataset construction and evaluation pipeline can be found in Section~\ref{?}.
\input{table/ocr_lumina}

% \subsubsection{Generation training improves recognition but not the reverse.} Table~\ref{tab:ocr} summarizes the OCR transfer results. A clear asymmetry emerges between the two transfer directions. Training with the generation objective alone leads to consistent improvement in OCR accuracy across models (\textit{gen}\ $\rightarrow$ \textit{und}). In contrast, training with the understanding objective yields limited gains in text rendering quality (\textit{und}\ $\rightarrow$ \textit{gen}). We attribute this asymmetry to the nature of the two tasks. This contrasts with the counting results in Section~\ref{sec:umms}, where bi-directional transfer was observed.
\subsubsection{OCR transfer is unidirectional: generation aids recognition, but not vice versa.} Table~\ref{tab:ocr} summarizes the results. A clear asymmetry emerges between the two transfer directions. Training with the generation objective alone leads to consistent improvement in OCR accuracy~(\textit{gen}~$\rightarrow$~\textit{und}), suggesting that learning to render text strengthens the model's character-level representations in ways that benefit recognition. In contrast, training with the understanding objective yields limited gains in text rendering quality~(\textit{und}~$\rightarrow$~\textit{gen}). This stands in contrast to the counting results in Section~\ref{sec:umms}, where transfer was bi-directional: for OCR, the benefit flows in only one direction. We hypothesize that this asymmetry comes from a fundamental difference in task demands. Generating legible text requires the model to internalize fine-grained glyph structure—stroke geometry, spacing, and character composition—which naturally supports recognition. Recognition, however, can rely on partial or contextual cues without fully encoding the visual details needed for accurate rendering, limiting its utility as a training signal for generation.
