\section{Discussion}
\subsection{(what determines transferability?)}

Taken together, our results across three tasks suggest that both the direction and the strength of transfer depend on the nature of the visual attributes a model must internalize to solve each task.


For counting and spatial relations, the key attributes are coarse semantic signals such as object numerosity and layout. These signals tend to be encoded at a similar level regardless of whether the model is trained with an understanding-oriented or generation-oriented objective. Notably, the semantic representations learned through understanding can also provide effective guidance for generation. This observation aligns with prior work showing that improving semantic representations can directly benefit generative performance~\cite{yu2024repa, zheng2025rae}. In our experiments, a model trained to understand counting develops representations that reliably capture the number of objects in an image, and these representations remain useful when the model is asked to generate images containing a specified number of objects. Similarly, training for spatial-relation understanding encourages representations that encode relative object configurations, which can be exploited during generation to produce the intended layout. In other words, semantic knowledge acquired through understanding can meaningfully condition the generation process. When the shared concept is sufficiently coarse and can be captured at a semantic level, this conditioning effect can operate in both directions, leading to bi-directional transfer.

In contrast, text recognition and text generation exhibit an asymmetric transfer pattern. A generation objective requires the model to synthesize individual characters with pixel-level fidelity, and the resulting fine-grained character representations transfer naturally to recognition. The reverse direction is comparatively weaker. While prior studies report that recognition training can facilitate text generation~\cite{ji2023improving, fang2025recognsyn, min2025tair}, we also observe only limited transfer from understanding to generation in this setting. A plausible explanation is that recognition-oriented training primarily emphasizes high-level textual content and identity, whereas text generation additionally demands low-level, spatially precise rendering details. Consequently, semantic knowledge extracted from recognition alone is insufficient to fully bridge the gap to pixel-accurate character synthesis. Overall, our findings indicate that when the shared concept requires fine-grained, pixel-level precision, semantic representations from understanding provide only partial support for generation.

\subsection{Transferability as a practical fine-tuning strategy}
\label{sec:finetune_strategy}
In the preceding sections, we observed that training on only one modality can yield improvements in the other, producing an effect similar to training both sides simultaneously. Here we discuss whether this property can be leveraged as a practical post-training strategy. In particular, our findings suggest that exploiting understanding-to-generation transfer can be practically advantageous.

\input{table/gen_fid}
\subsubsection{Robust transfer without generation quality loss.}

% In Sections~\ref{sec:umms} and~\ref{}, we observed that understanding training effectively injects task-relevant concepts into generation for both counting and spatial relations. For counting (Table~\ref{tab:gen_quality}), understanding-only and generation-only training achieve comparable task accuracy, yet their impact on image quality diverges sharply. Understanding training (\textit{und}$\rightarrow$\textit{gen}) improves both FID and IS over the baseline, whereas generation training (\textit{gen}$\rightarrow$\textit{gen}) degrades both metrics substantially. For spatial relations (Table~\ref{tab:spatial_position_transfer}), generation training yields numerically higher accuracy (81.25\% vs.\ 73.75\%), but direct generation supervision on synthetic data causes the model to overfit to the synthetic distribution, resulting in a loss of visual realism and diversity (Figure~\ref{fig:spatial_pos}).
\input{fig/error_decomp}
In Sections~\ref{sec:umms} and~\ref{}, we observed that understanding training effectively injects task-relevant concepts into generation for both counting and spatial relations. For counting (Table~\ref{tab:gen_quality}), understanding-only and generation-only training achieve comparable task accuracy, yet their impact on image quality diverges sharply. Understanding training (\textit{und}$\rightarrow$ \textit{gen}) improves both FID and IS over the baseline, whereas generation training (\textit{gen}$\rightarrow$
\textit{gen}) degrades both metrics substantially. For spatial relations (Table~\ref{tab
:spatial_position_transfer}), generation training yields numerically higher accuracy (81.25\% vs.\ 73.75\%), but direct generation supervision on synthetic data causes the model to overfit to the synthetic distribution, resulting in a loss of visual realism and diversity (Figure~\ref{fig:spatial_pos}). Figure~\ref{fig:error_decomp} shows the error decomposition, indicating that understanding training primarily reduces position errors while detection failures remain largely unchanged, suggesting that the model's generation behavior remains close to the original distribution.
In many practical scenarios, synthetic data may be the only available supervision for generation. However, without careful control over the data domain, post-training directly on such data risks distribution collapse. Understanding training sidesteps this problem entirely. Since the model never optimizes pixel-level generation on out-of-distribution images, generative diversity is preserved while the target concept still improves meaningfully through transfer.

\input{table/general_bench}
\subsubsection{Preserving general capabilities.}
For transfer to serve as a viable post-training strategy, task-specific understanding training must not degrade the model's existing capabilities. To verify this, we evaluate the model trained on counting understanding using general multimodal understanding benchmarks including POPE, MME, and MMBench (Table~\ref{tab:general_bench}. The results show that performance is maintained or slightly improved compared to the baseline, confirming that task-specific understanding training does not erode the model's broader knowledge.

\vspace{0.5em}
Taken together, these cases illustrate a general principle. Understanding training acts as a semantic regularizer that steers generation toward the target capability without corrupting the learned generative distribution. Our findings suggest that transferability is not merely an analytical observation but a viable post-training strategy.
% The preceding analysis characterizes \emph{when} transfer occurs. We now argue that transfer can be deliberately exploited: when the goal is to improve a specific generation capability, training via understanding and relying on transfer (\textit{und}\ $\rightarrow$ \textit{gen}) can be preferable to direct generation fine-tuning.

% \input{table/gen_fid}
% \paragraph{Counting: \textit{und}\ training is Pareto-dominant.}
% Table~\ref{tab:gen_quality} compares Lumina-DiMOO's counting generation under three conditions: no fine-tuning (Baseline), understanding-only training (\textit{und}\ $\rightarrow$ \textit{gen}), and generation-only training (\textit{gen}\ $\rightarrow$ \textit{gen}). Both fine-tuned conditions achieve the same task accuracy (0.57), and \textit{und}\ $\rightarrow$ \textit{gen}\ yields a lower MAD. More strikingly, image quality diverges: \textit{und}\ $\rightarrow$ \textit{gen}\ improves both Inception Score (IS) and FID over the baseline, whereas \textit{gen}\ $\rightarrow$ \textit{gen}\ degrades both metrics substantially. Understanding training thus provides the same task-level improvement at no cost to—and even a benefit for—perceptual quality, making it Pareto-dominant over direct generation training.

% \input{fig/error_decomp}
% \paragraph{Spatial position: \textit{und}\ training as a safer strategy.}
% For spatial position (Table~\ref{tab:spatial_position_transfer}), \textit{gen}\ $\rightarrow$ \textit{gen}\ is numerically stronger (81.25\% vs.\ 73.75\%). Figure~\ref{fig:error_decomp} decomposes the errors into detection failures (object not found) and position errors (object detected but misplaced). Generation training reduces both error types, whereas understanding training primarily lowers position errors while detection failures decrease only modestly. However, because the training data is synthetic (composited ImageNet crops), direct generation supervision causes the model to overfit to the synthetic distribution—generated images lose visual realism and diversity (see Figure~\ref{fig:spatial_pos}). In many practical scenarios, synthetic data is the only available supervision for generation, and training directly on it risks distribution collapse. Understanding training sidesteps this problem entirely: the model never optimizes pixel-level generation on synthetic images, so generative diversity is preserved while spatial reasoning still improves meaningfully via transfer.

% Together, these two cases illustrate a general principle: understanding training acts as a \emph{semantic regularizer} that steers generation toward the target capability without corrupting the learned generative distribution. Transferability is therefore not merely an analytical finding—it is a practical fine-tuning strategy.

\subsection{Joint training is not the answer}
\label{sec:joint_training}
\input{fig/joint_training}

% A natural response to the above observation is to ask: why not train on both objectives simultaneously? We find that joint \textit{und}+\textit{gen}\ training introduces \emph{seesaw interference} between the two objectives. Figure~\ref{fig:joint_training} shows Lumina-DiMOO trained with both counting understanding and generation losses over 30 epochs. During the first ${\sim}$15 epochs, understanding accuracy rises while generation accuracy declines; during the latter half, generation recovers at the expense of understanding. The two objectives oscillate in opposition, competing for representational capacity even within a low-rank LoRA adapter.

% This observation is consistent with prior analyses showing that understanding and generation demand distinct alignment patterns across network depth~\cite{li2025unifork}, and that joint optimization can lead to representational compromise. The seesaw behavior further motivates the sequential transfer strategy of Section~\ref{sec:finetune_strategy}: train on one objective, then rely on transfer for the other. This avoids representational conflict and yields stable improvements in both capabilities.

Throughout this work, we have explored transferability in UMMs and confirmed that for certain tasks, knowledge learned through one objective can boost performance in the other direction. 

Throughout this work, we examined whether knowledge learned under one objective in unified multimodal models can transfer to the other. Our experiments show that such transfer can occur. For several tasks, training with one objective improves performance when evaluated under the other. This naturally raises the question of whether training both objectives simultaneously leads to improved performance for both.

To investigate this, we trained Lumina-DiMOO on counting understanding and generation jointly. As shown in Figure~\ref{fig:joint_training}, the two objectives do not improve together during training. Instead, their performance exhibits an oscillatory pattern. Understanding accuracy increases in the first half of training while generation accuracy declines, and the trend reverses in the latter half. This behavior suggests that the presence of transfer between objectives does not necessarily translate into consistent gains when the two are optimized simultaneously.

One possible explanation is that understanding and generation require different alignment patterns within the network, which may produce conflicting optimization signals during joint training~\cite{li2025unifork}. At the same time, the specifics of our experimental setting may also play a role. Our study relies on post-training with tightly paired and concept-focused data, which differs from the heterogeneous data distributions used in large-scale multimodal pre-training. Such a training regime may make joint optimization more difficult. Understanding under what conditions joint training leads to complementary improvements remains an open question.

\subsection{Importance of regularization: LoRA vs.\ full fine-tuning}
\label{sec:lora_vs_full}

The transfer results reported throughout this paper rely on LoRA fine-tuning. We now show that this choice of regularization is not incidental—it is critical. In an additional experiment with Janus-Pro, we compare LoRA fine-tuning (rank 128) against full fine-tuning of the transformer backbone (with the language modeling head frozen). Under full fine-tuning, generation collapses entirely: the model produces no coherent images despite the generation head remaining untouched. LoRA fine-tuning, by contrast, yields stable training and the transfer patterns reported in Table~\ref{tab:transfer}.

The explanation is straightforward. LoRA constrains parameter updates to a low-rank subspace, acting as implicit regularization that preserves the pretrained generative representations. Full fine-tuning permits unconstrained parameter drift across the entire backbone, catastrophically disrupting the learned visual-generative mappings. For UMMs, regularization is not optional—it determines whether the model retains its multimodal capabilities at all.

